### 2024-11-24 other/EDA.qmd
added many exploratory graphs

---
title: "EDA"
author: 
  - Robert Ford
format: pdf
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(scales)
library(knitr)
library(kableExtra)

df <- read_csv(here::here("data/analysis_data/augment_data.csv"))

df <- df %>% select(-link)

```

```{r fig-template}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

```

```{r fig-head}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

kable(head(df, 5), caption = "Sample of the Data")
```

This graph shows how much was awarded by month
```{r fig-award-amount-month}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

df <- df %>%
  mutate(year_month = floor_date(award_date, "month"))

# Aggregate the amount by 'year_month'
monthly_spending <- df %>%
  group_by(year_month) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE))

# Plot the line graph with formatted y-axis labels
ggplot(monthly_spending, aes(x = year_month, y = total_amount)) +
  geom_line(color = "blue") +
  labs(title = "Amount Spent Each Month",
       x = "Time",
       y = "Amount Awarded") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r fig-template4}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

df$amount <- as.numeric(df$amount)

# Ensure 'award_date' is in Date format (if not already)
# df$award_date <- as.Date(df$award_date, format="%Y-%m-%d")  # Commented out since it's already a Date

# Identify the top 5 buyers by total amount awarded
top_buyers <- df %>%
  group_by(buyer) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE)) %>%
  arrange(desc(total_amount)) %>%
  slice_head(n = 3) %>%
  pull(buyer)

# Filter the data for only the top 5 buyers
df_top_buyers <- df %>%
  filter(buyer %in% top_buyers)

# Create a 'year_month' column using lubridate
df_top_buyers <- df_top_buyers %>%
  mutate(year_month = floor_date(award_date, "month"))

# Aggregate the amount by 'year_month' and 'buyer'
monthly_spending <- df_top_buyers %>%
  group_by(year_month, buyer) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE), .groups = 'drop')

# Plot the line graph
ggplot(monthly_spending, aes(x = year_month, y = total_amount, color = buyer, group = buyer)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Amount Spent Each Month by Top 5 Buyers",
       x = "Time",
       y = "Amount Awarded",
       color = "Buyer") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r fig-template3}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

df$amount <- as.numeric(df$amount)

# Summarize total amount spent by each buyer
buyer_totals <- df %>%
  group_by(buyer) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE)) %>%
  arrange(desc(total_amount))

# Select the top ten buyers
top_ten_buyers <- buyer_totals %>%
  slice_head(n = 10)

# Plot the bar chart
ggplot(top_ten_buyers, aes(x = reorder(buyer, total_amount), y = total_amount)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Top 10 Buyers by Total Amount Spent",
       x = "Buyer",
       y = "Total Amount Spent") +
  scale_y_continuous(labels = comma) +
  theme_minimal()

```

```{r fig-template2}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

supplier_totals <- df %>%
  group_by(supplier) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE)) %>%
  arrange(desc(total_amount))

# Select the top ten suppliers
top_suppliers <- supplier_totals %>%
  slice_head(n = 10)

# Wrap supplier names for better legibility
top_suppliers$supplier <- str_wrap(top_suppliers$supplier, width = 15)

# Plot the bar chart
ggplot(top_suppliers, aes(x = reorder(supplier, -total_amount), y = total_amount)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Amount Awarded to Top 10 Suppliers",
       x = "Supplier",
       y = "Total Amount Awarded") +
  scale_y_continuous(labels = comma) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.margin = unit(c(1, 1, 1, 2), "cm")  # Adjust margins if necessary
  )

```

```{r fig-template1}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

df <- df %>%
  mutate(contract_length_days = as.numeric(difftime(end_date, start_date, units = "days")))

# Group by buyer and calculate average length of contract and average contract size
buyer_stats <- df %>%
  group_by(buyer) %>%
  summarise(
    `Average Length of Contract (Days)` = mean(contract_length_days, na.rm = TRUE),
    `Average Contract Size` = mean(amount, na.rm = TRUE)
  )

# Reshape the data to have buyers as columns and metrics as rows
buyer_stats_long <- buyer_stats %>%
  pivot_longer(cols = -buyer, 
               names_to = "Metric", 
               values_to = "Value")

# Now pivot wider to have buyers as columns
buyer_stats_wide <- buyer_stats_long %>%
  pivot_wider(names_from = buyer, values_from = Value)

# Optionally, format the numbers
buyer_stats_wide <- buyer_stats_wide %>%
  mutate(across(-Metric, ~ round(.x, 2)))

# Display the table using knitr::kable
kable(buyer_stats_wide, caption = "Average Length of Contract and Average Contract Size by Buyer")

```